Learning constraints in tabular data

### 2016-05-10 Paper/ECAI/tabular.tex
quick and dirty update to CIKM/ACM style
some reworking of problem statement, not finished at all

\documentclass{ecai}
\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[numbers]{natbib}
\usepackage{notoccite}
\usepackage{framed}
\usepackage{amsmath}
\usepackage{tabularx}

@@

\usepackage{amsthm}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

@@

\title{Tabular Constraint Learning}

\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{----------------------} }

@@

\section{Problem Statement}
A {\em constraint satisfaction problem} (CSP) consists of a set of constraints and the
task is to find assignments of values to the variables in the CSP so that all constraints are satisfied.
In the context of spreadsheets the variables typically refer to cells in tables, and solving the corresponding CSP problem means completing the tables so that all constraints in the spreadsheet are satisfied.

While the traditional CSP problem is given a set of variables and a set of constraints, and the task is to find an assignment, the {\bf inverse CSP} or the {\bf Tabular Constraint Learning Problem} can be defined as follows.\\
\begin{definition}[Tabular Constraint Learning]\label{def:problem_statement}
  Let ${\cal T}$ a set of instantiated tables (all cells have values),  ${\cal C}$ be a set of {\template}s, ${\cal G}$ be a set of groups  for the tables ${\cal T}$, then find the set of all constraints $s(G_1', ..., G_n')$ that are satisfied  in ${\cal T}$, for which all $G_i' \subseteq G_i \in {\cal G}$, and the groups $G_1', ... , G_n'$ satisfy the signature of the constraint template~$s \in {\cal C}$.
\end{definition}
% {\bf Given }
% \begin{enumerate}
% \item
% a set of instantiated tables (so all cells have values) ${\cal T}$;
% \item
% a set of {\template}s ${\cal C}$;
% \item
% a set of groups ${\cal G}$ for the tables ${\cal T}$;
% \end{enumerate}
% \noindent
% {\bf Find}  the set of all constraints $s(G_1', ..., G_n')$ that 
% \begin{enumerate}
% \item are satisfied  in ${\cal T}$;
% \item for which all $G_i' \subseteq G_i \in {\cal G}$ ;
% \item the groups $G_1', ... , G_n'$ satisfy the signature of the constraint template~$s \in {\cal C}$.
% \end{enumerate}
For instance, Figure 3 contains the solution to the tabular constraint learning problem
that uses the tables of Figure 1, the signatures of Table 1 and the set of groups mentioned in (\samuel{show groups somewhere}).

% \luc{Delete from here}
% \luc{ADD/REFER TO EXAMPLE}

@@

The key observation here is that essentially the problem is a constraint enumeration problem, where each constraint is independent. This property comes from fact that in spreadsheets each formula is applied based on the existing data in the cells. This allows learning constraints independently of each other by examining constraint satisfaction on the groups.

@@

\newcommand{\tcl}{Tabular Constraint Learning}
\newcommand{\ctl}{Condensed Tabular Constraint Learning}
\section{Approach to Tabular Constraint Learning}
% \sergey{Luc wants this to be more intuitive, should we just add some examples here to make it easier to follow and refer to the running example?}
This section will describe our approach to \tcl.
Our algorithm reasons on the level of groups and vectors on the data and assumes that these are provided.
Often, however, this not the case and a possible approach is to extract the groups from the CSV file directly.
This requires discovering \textit{tables} and \textit{groups}.
Our implementation offers automated procedures for both steps.

@@

\subsubsection{Generality ordering}
\sergey{should we show all the formulas here? or some maybe and say that the specificity DAG can be extracted?}

@@

\tias{Move this to problem statement second part}
A constraint~$c_1$ is a special case of another constraint~$c_2$ iff whenever $c_1$ is true, $c_2$ must also be true.
We consider $c_1$ to be more specific than $c_2$ or $c_2$ to be more general than $c_1$.
A constraint template~$s_1$ is more specific than a template~$s_2$ iff a constraint applying $s_1$ to a set of subgroups is always more specific than a constraint applying $s_2$ to (part of) the same set.
\samuel{Formalize: $\forall x1..x_k s_1(x_1...x_n) => s_2(x_1...x_k)$ :: looking for most specific}

>>>>

\documentclass{sig-alternate-05-2015}

\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[numbers]{natbib}
\usepackage{notoccite}
\usepackage{framed}
\usepackage{amsmath}
\usepackage{tabularx}

@@

%\usepackage{amsthm} % incompatible with ACM
%\theoremstyle{definition}
%\newtheorem{definition}{Definition}
\newdef{definition}{Definition} % ACM specific

@@

\title{Learning constraints in tabular data}
% see http://www.acm.org/binaries/content/assets/publications/article-templates/sig-alternate-sample.tex

%\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{----------------------} }

@@

\section{Problem Statement}
The problem of learning constraints from tabular data can be seen as an inverse {\em constraint satisfaction problem} (CSP). A CSP consists of a set of constraints and the
task is to find assignments of values to the variables in the CSP so that all constraints are satisfied.
In the context of spreadsheets where variables could represent cells in tables, solving a corresponding CSP problem would mean completing the tables so that all constraints of the spreadsheet are satisfied.

The inverse problem is, given an assignment to all cells, to find the constraints that are present in the spreadsheet.
We define the {\bf Tabular Constraint Learning Problem} as follows:
%
\begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}
  Given ${\cal G}$ a set of groups derived from fully instantiated tables ${\cal T}$ (all cells have values) and a set of {\template}s ${\cal C}$ ; find the set of all constraints $s(G_1', ..., G_n')$ that match the syntax of a \template in ${\cal C}$ and where the groups $G_1', ... , G_n'$ with $\forall i, G_i' \subseteq G_i, G_i \in {\cal G}$ satisfy both the signature and the definition of that \template.
\end{definition}
% {\bf Given }
% \begin{enumerate}
% \item
% a set of instantiated tables (so all cells have values) ${\cal T}$;
% \item
% a set of {\template}s ${\cal C}$;
% \item
% a set of groups ${\cal G}$ for the tables ${\cal T}$;
% \end{enumerate}
% \noindent
% {\bf Find}  the set of all constraints $s(G_1', ..., G_n')$ that 
% \begin{enumerate}
% \item are satisfied  in ${\cal T}$;
% \item for which all $G_i' \subseteq G_i \in {\cal G}$ ;
% \item the groups $G_1', ... , G_n'$ satisfy the signature of the constraint template~$s \in {\cal C}$.
% \end{enumerate}
An example solution for the type-consistent groups of the tables in Figure~\ref{fig:main_example} and {\template}'s in Table~\ref{table:constraints} is shown in Figure~\ref{fig:sol_example}.

@@

As is well-studied in Inductive Logic Programming, some constraints can \textit{entail} other constraints, meaning that some constraints are \textit{implied} by others. More formally, a constraint $c_1$ implies another constraint~$c_2$ iff whenever $c_1$ is true, $c_2$ must also be true. As a result, a solution to the tabular constraint learning problem may contain constraints that are implied by others. We can hence say that there is \textit{redundancy} in the solution set of constraints.

Two types of impliciation:

1) on constraint definitions for same signature (example: permutation and alldiff or foreign-key and alldiff)

2) on constraint signature for same definition (example: ?) \textit{Formally, we say that a constraint template $t_1$ is \textit{more specific} than a constraint template $t_2$ iff for $t_1$ and $t_2$  holds the statements of the form: $\forall g_1, \dots, g_n{:}~ t_1(g_1,\dots,g_n) \rightarrow t_2(g_h, \dots, g_k)$ where $1 \leq h \leq k \leq n$. The intuition behind the formula is the following, if a list of groups is a solution for $t_1$ then its subset is also a solution for $t_2$.}

3) combination

\tias{OLD TEXT:}

A constraint template~$s_1$ is more specific than a template~$s_2$ iff a constraint applying $s_1$ to a set of subgroups is always more specific than a constraint applying $s_2$ to (part of) the same set.
\samuel{Formalize: $\forall x1..x_k s_1(x_1...x_n) => s_2(x_1...x_k)$ :: looking for most specific}

@@

\newcommand{\tcl}{Tabular Constraint Learning}
\newcommand{\ctl}{Condensed Tabular Constraint Learning}
\section{Approach to Tabular Constraint Learning}
% \sergey{Luc wants this to be more intuitive, should we just add some examples here to make it easier to follow and refer to the running example?}
This section will describe our approach to \tcl.

The key observation is that essentially the problem is a constraint enumeration problem, where each constraint is independent. This property comes from fact that in spreadsheets each formula is applied based on the existing data in the cells. This allows learning constraints independently of each other by examining constraint satisfaction on the groups.

Our algorithm reasons on the level of groups and vectors on the data and assumes that these are provided.
Often, however, this not the case and a possible approach is to extract the groups from the CSV file directly.
This requires discovering \textit{tables} and \textit{groups}.
Our implementation offers automated procedures for both steps.

@@

\subsubsection{Generality ordering}
\tias{ordering is because some constraint templates use other templates in their signature, doesn't have to be more complicated?}

@@
